\( \tool \): JavaScript IR-based Semantics Extraction Toolchain

### 2020-04-30 ase20/eval.tex
Till Sec. 5.3

    BigInt-1 &
    Expression &
    \makecell[l]{Using wrong variable \code{oldvalue} instead of \\
    \code{oldValue} in {\bf Evaluation} of {\it UpdateExpression}.} &
    \textsf{X} &
    2019-09-27 &
    2020-04-23 &
    209 days &
    \inred{XX} \\\hline

    BigInt-2 &
    Number &
    \makecell[l]{Using {\bf ToInt32} instead of {\bf ToUint32} \\ in {\bf
    Number::unsignedRightShift}.} &
    \textsf{X} &
    2019-09-27 &
    2020-04-23 &
    209 days &
    \inred{XX} \\\hline

    BigInt-3 &
    Number &
    \makecell[l]{Using {\bf ToUint32} instead of {\bf ToInt32} in {\bf
    NumberBitwiseOp}.} &
    \textsf{X} &
    2019-09-27 &
    2020-04-23 &
    209 days &
    \inred{XX} \\\hline

    BigInt-4 &
    Number &
    \makecell[l]{Not handling BigInt values in {\bf Number} constructor.} &
    \textsf{O} &
    2019-09-27 &
    2019-11-19 &
    53 days &
    \inred{XX}
  \end{tabular}
\end{table*}

@@

\subsection{Forward Compatibility}

We show the forward compatibility of \( \tool \) by applying it to the proposals
that are ready for inclusion in the next ECMAScript (ES11, 2020). Because
ECMAScript administered as an open-source project, various proposals for new
language features are available with their own specification changes and tests.
A separate repository~\cite{proposals} maintains such proposals with six
different stages: Stage 0 to Stage 3, Finished, and Inactive.  Each proposal
starts with Stage 0, and the TC39 committee regularly examines Stage 3 proposals
to decide their next stages.  If a proposal is confirmed, the committee changes
its stage to Finished and integrates it into the next version of ECMAScript.
Otherwise, the committee changes its stage to Inactive.  Among them, we applied
\( \tool \) to all Finished proposals decided to be included in ES11.

\begin{table}[t]
  \centering
  \caption{Evaluation result of \( \tool \) for Finished stage proposals that
  will be included in ES11.}
  \label{table:spec-prop-result}
  \vspace*{-0.5em}
  \small
  \begin{tabular}{c?r|r?r?r?r}
    \hline
    \multicolumn{1}{c?}{\multirow{2}{*}{\bf Name}} &
    \multicolumn{2}{c?}{\bf \( \Delta \) \# Prod.} &
    \multicolumn{1}{c?}{\multirow{2}{*}{\bf \( \Delta \) \# Steps}} &
    \multicolumn{1}{c?}{\multirow{2}{*}{\bf \( \Delta \) \# Tests}} &
    \multicolumn{1}{c}{\multirow{2}{*}{\bf \# Tests}} \\\cline{2-3}

    &
    \multicolumn{1}{c|}{\bf Lex.} &
    \multicolumn{1}{c?}{\bf Syn.} &
    &\\\hline\hline

    \code{matchAll} of \code{String} &
    0 &
    0 &
    9/9 &
    5/5 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

    \code{import()} &
    0 &
    2 &
    22/23 &
    0/0 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

    \code{BigInt} &
    4 &
    0 &
    258/323 &
    \inred{207}/207 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

    \code{Promise.allSettled} &
    0 &
    0 &
    79/85 &
    50/50 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

    \code{globalThis} &
    0 &
    0 &
    1/1 &
    1/1 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

    \code{for-in} mechanics &
    0 &
    0 &
    36/37 &
    0/0 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

    Optional Chaining &
    3 &
    3 &
    72/74 &
    19/19 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

    Nullish Coalescing Op. &
    1 &
    4 &
    10/10 &
    21/21 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

    \code{import.meta} &
    0 &
    2 &
    15/15 &
    0/0 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

    {\bf Total} &
    8 &
    11 &
    502/577\\\cline{1-4}
  \end{tabular}
\end{table}

Table~\ref{table:spec-prop-result} shows the result of applying \( \tool \) to 9
Finished proposals for ES11.  For all proposals, \inred{XX} lexical and
\inred{XX} syntactic productions are modified and \( \tool \) successfully
updated the changes to the synthesized JavaScript parser.  The updated parser
parses all applicable tests for all proposals.  For abstract algorithms,
\inred{XXX} steps out of \inred{XXX} steps are automatically converted to the
corresponding \( \ires \) instructions by the algorithm compiler without
changing of the compile rules.  Thus, our tool has \inred{XX.XX\%} success rate
on average for future proposals.  Similar with
Section~\ref{sec:check-with-tests}, we manually implemented missing part of the
AST-IR translator for each Finished proposal and checked each extracted
semantics with official tests.  Each of them passed all applicable tests except
the semantics for \code{BigInt} proposal.  It fails \inred{XX} tests out of 207
applicable tests for the proposal and \inred{XXXX} tests out of \inred{XXXXX}
applicable tests for ES10.

Based on the failed tests, we found three undiscovered specification errors
(BigInt-1, BigInt-2, and BigInt-3), and one reported error (BigInt-4) in the
BigInt proposal as described in Table~\ref{table:spec-errors}.  All of
them were confirmed by TC39 and will be fixed in the next release.  The BigInt
proposal added a new type of primitives BigInt and unified the type with
original Number type to Numeric types.  Thus, it not only added new
algorithms for BigInt but also modified all existing algorithms related to
number values.  The error BigInt-1 is just a simple misuse of the variable
\code{oldValue} in {\bf Evaluation} of {\it UpdateExpression}. BigInt-2 and
BigInt-3 break the backward compatibility because of the incorrectly mixed uses
of {\bf ToInt32} and {\bf ToUint32} algorithms in unsigned right shift operators
and bitwise operators, respectively. The last one BigInt-4 arises because the
Number constructor does not consider BigInt primitives.  On average, ECMAScript
included these four specification errors in \inred{XXX} days.

After fixing all specification errors, we extracted the semantics for the BigInt
proposal again.  The updated semantics passed all 207 applicable tests for the
proposal and \inred{XXXXX} applicable tests for ES10.  Thus, \( \tool \) also
correctly extract IR-based semantics from future proposals.

>>>>

    BigInt-1 &
    Expression &
    \makecell[l]{Using the wrong variable \code{oldvalue} instead of
    \code{oldValue} in \\ {\bf Evaluation} of {\it UpdateExpression}} &
    \textsf{X} &
    2019-09-27 &
    2020-04-23 &
    209 days &
    \inred{XX} \\\hline

    BigInt-2 &
    Number &
    \makecell[l]{Using {\bf ToInt32} instead of {\bf ToUint32} in \\ {\bf
    Number::unsignedRightShift}} &
    \textsf{X} &
    2019-09-27 &
    2020-04-23 &
    209 days &
    \inred{XX} \\\hline

    BigInt-3 &
    Number &
    \makecell[l]{Using {\bf ToUint32} instead of {\bf ToInt32} in {\bf
    NumberBitwiseOp}} &
    \textsf{X} &
    2019-09-27 &
    2020-04-23 &
    209 days &
    \inred{XX} \\\hline

    BigInt-4 &
    Number &
    \makecell[l]{Unhandling BigInt values in the {\bf Number} constructor} &
    \textsf{O} &
    2019-09-27 &
    2019-11-19 &
    53 days &
    \inred{XX}
  \end{tabular}
\end{table*}

@@

\begin{table}[t]
  \centering
  \caption{Proposals that will be included in ES11}
  \label{table:spec-prop-result}
  \vspace*{-0.5em}
  \small
  \begin{tabular}{@{}c@{}?r|@{~}r@{~}|@{}c@{}|@{}c@{}|c@{}}
    \multicolumn{1}{@{}c@{}?}{\multirow{2}{*}{\bf Proposal}} &
    \multicolumn{2}{@{~}c@{~}|}{\bf \( \Delta \) \# Prod.} &
    \multicolumn{1}{@{~}c@{~}|}{\multirow{2}{*}{\bf \( \Delta \)\! \#\! Steps}} &
    \multicolumn{1}{@{~}c@{~}|}{\multirow{2}{*}{\bf \( \Delta \)\! \#\! Tests}} &
    \multicolumn{1}{c@{}}{\multirow{2}{*}{\bf \# Tests}} \\\cline{2-3}

    &
    \multicolumn{1}{@{~}c@{~}|}{\bf Lex.} &
    \multicolumn{1}{@{~}c@{~}|}{\bf Syn.} &
    &\\\toprule

\footnotesize
    \scode{matchAll} of \scode{String} &
    0 &
    0 &
    9/9 &
    5/5 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

    \scode{import()} &
    0 &
    2 &
    22/23 &
    0/0 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

    \scode{BigInt} &
    4 &
    0 &
    258/323 &
    \inred{207}/207 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

    \scode{Promise.allSettled } &
    0 &
    0 &
    79/85 &
    50/50 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

    \scode{globalThis} &
    0 &
    0 &
    1/1 &
    1/1 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

\footnotesize
    \scode{for-in}  mechanics &
    0 &
    0 &
    36/37 &
    0/0 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

\footnotesize    Optional Chaining &
    3 &
    3 &
    72/74 &
    19/19 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

\footnotesize   Nullish Coalescing Op. &
    1 &
    4 &
    10/10 &
    21/21 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

    \scode{import.meta} &
    0 &
    2 &
    15/15 &
    0/0 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

    {\bf Total} &
    8 &
    11 &
    502/577 &
  \end{tabular}
\end{table}

\subsection{Forward Compatibility}
We evaluated whether \( \tool \) is forward compatible by applying it
to the proposals that are ready for inclusion in ES11.  Because
ECMAScript is an open-source project, various proposals for new
language features are available with their own specification changes and tests.
A separate repository~\cite{proposals} maintains such proposals with six
stages: Stage 0 to Stage 3, Finished, and Inactive.  A proposal
starts with Stage 0, and the TC39 committee examines proposals in Stage 3.
If a proposal is confirmed, the committee changes
its stage to Finished and integrates it into the next ECMAScript.
Otherwise, the committee changes its stage to Inactive.
Thus, we applied \( \tool \) to all nine proposals in the Finished stage.

Table~\ref{table:spec-prop-result} shows the result of applying \( \tool \) to the
Finished proposals.  Collectively, the proposals modified \inred{XX} lexical and
\inred{XX} syntactic productions, and \( \tool \) successfully
synthesized parsers for them.  The newly synthesized parsers parse all
applicable tests for all proposals.  For abstract algorithms,
\inred{XXX} steps out of \inred{XXX} are automatically converted to
corresponding \( \ires \) instructions by \textsf{Algorithm Compiler}
without changing {\sf Compile Rules}.  Thus, \( \tool \) has the
success rate of \inred{XX.XX\%} on average for forthcoming proposals.

We checked the extracted semantics from the proposals
by implemented missing parts of the AST-IR translator for
each proposal and checking the semantics with Test262.  All of them
passed all applicable tests except the semantics from the
\code{BigInt} proposal.  It failed for \inred{XX} tests out of 207
applicable tests for the proposal and \inred{XXXX} tests out of
\inred{XXXXX} applicable tests for ES10.

Using the failed tests, we discovered four specification errors in the
BigInt proposal: three new errors (BigInt-1 to BigInt-3) and one known
error (BigInt-4) as summarized in Table~\ref{table:spec-errors}.
All of them were confirmed by TC39 and will be fixed in the next ECMAScript.
The proposal added two new types: BigInt as a new type of primitives
and Numeric as a unified type of the original Number type and the new BigInt type.
Therefore, it not only added new algorithms for BigInt but also modified
all existing algorithms for number values.
The error BigInt-1 is due to a simple misuse of the variable 
\code{oldValue} in {\bf Evaluation} of {\it UpdateExpression}.
BigInt-2 and BigInt-3 break the backward compatibility because of
the incorrectly mixed uses of {\bf ToInt32} and {\bf ToUint32}
in unsigned right shift operators and bitwise operators, respectively.
Finally, BigInt-4 happened because the proposal did not consider
BigInt primitives in the Number constructor.  On average, four
specification errors existed for \inred{XXX} days in the proposal.

After fixing the errors in the proposal, we extracted a semantics from
the revised specification.  The extracted semantics passed all 207
applicable tests for the proposal and \inred{XXXXX} applicable tests
for ES10.  Thus, \( \tool \) also correctly extracts an IR-based
semantics from future proposals, which implies that it is forward compatible.
